### Week 5: Dimensionality reduction techniques

```{r}
date() 
```

## 1) Data wrangling exercises 
R-Script:
https://github.com/venkatharina/IODS-project/blob/Data/create_human.R

## 2) Analysis exercises

```{r}
library(dplyr)
library(readr)
library(corrplot)
library(tibble)
library(GGally)

#reading data in
human <- read_csv("https://raw.githubusercontent.com/KimmoVehkalahti/Helsinki-Open-Data-Science/master/datasets/human2.csv")

#moving the country names to rownames
human_ <- column_to_rownames(human, "Country")

#visualisation of data:
ggpairs(human_, progress = FALSE)
#summary of data:
summary(human_)
#correlation matrix and its visual representation:
cor(human_)
corrplot(cor(human_))
#description of data:
#Seems that Perli.F and LaboFM has least correlations with other variables
#Edu.FM2 seem to positively correlate with Life.Exp, Edu.Exp and GNI and 
#negatively correlate with Mat.Mor and Ado.Birth
#Life.Exp positively correlates with Edu2.FM, Edu.Exp and GNI and 
#negatively correlates to Mat.Mor and Ado.Birth
#Overall higher education seems to decrease adolescent birth rate and 
#maternal mortality

#Principal component analysis (PCA) with biplot
pca_human <- prcomp(human_)
pca_human
biplot(pca_human, choices = 1:2)

#Principal component analysis (PCA) with biplot with standardized data
human_std <- scale(human_)
pca_human2 <- prcomp(human_std)
pca_human2
biplot(pca_human2, choices = 1:2)
#standardized model is much more readable than non-standardized
#non-standardized:
#GNI seems higher in countries like Singapore, Norway
#From the rest, it is hard to find out

#standardized:
#If we see from the center of the graph:
#Up we seem more females in parliament (Parli.F) and Labour force of females (Labo.FM) (counties like Bolivia, Tanzania)
#Left we see more Education expectation, GNI, educated females, life expectancy (countries like Singapore, Ireland, Canada)
#On right we see maternity mortality and adolescents birth rates (countries like Sierra Leone, Gambia, Burkina Faso)

#Personal opinion:
#In more female educational countries, GNI and life expectancy seems higher and adolescent birth rate and maternal mortality rates are lower
#And males and females have equal labor force, and equal parliament
#Many Western Countries seem to be in this group

#300 individuals were asked how they drink tea (18 questions) and what are their product's perception (12 questions). 
#In addition, some personal details were asked (4 questions).
#loading data:
tea <- read.csv("https://raw.githubusercontent.com/KimmoVehkalahti/Helsinki-Open-Data-Science/master/datasets/tea.csv", stringsAsFactors = TRUE)
#visualisation of data:
view(tea)
str(tea)
dim(tea) #300rows and 36variables
#pivot_longer(tea, cols = everything()) %>% 
#  ggplot(aes(value)) + facet_wrap("name", scales = "free")
#MCA analysis:
#library("FactoMineR")
#install.packages("FactoMineR")

#URL: 'https://cran.rstudio.com/src/contrib/FactoMineR_2.9.tar.gz'

#install.packages(
#  "https://cran.r-project.org/src/contrib/Archive/FactoMineR/FactoMineR_2.7.tar.gz", 
#  repos = NULL, type = "source"
#)


